

\chapter{Acknowledgements} % this needs to be before \mainmatter.

I've encountered a lot of people over the past few years that have helped me, either directly or indirectly, and I want to thank you all.  I hope I don't forget anyone...but if I do, know that it's not on purpose.

Thanks, most of all, to my wife Shannon.  You were here when I started this, and you're still here at the end of it.  That's dedication!  I see every day the things you do to support me.  

I also wouldn't have succeeded in finishing this marathon without the help and support of my family, especially my parents.  You've been there to help and advise me since day one, and are still there helping.  I couldn't have asked for more.

By the time thesis writing comes around, the first and second year of graduate school pain seems like a distant memory.  But that was shared suffering (Jackson E\&M comes to mind...) and I'm grateful for your help, friendship, and fun that we had, Dave, Anya, Evan, Anthony, Alexey, Hannah, Ryan.  It would have been impossible to stay sane and survive that without you.

I've had the pleasure of working with and knowing a lot of people over the years, and one way or another you've been really helpful as I worked towards this thesis.  Thanks to Kyungeun, Sue and Lil, David K. and David C., Kazu, Mike M., Joseph, Kirby, Tingjun, Zarko, Georgia, Dave S., ... and everyone else that I've missed.

Josh and Mitch, your advice, feedback, and conversations really steered me in the right direction over the past year and turned this into an analysis I'm proud of.  Thanks.

Ariana, Elena, Brooke, Xiao, and Serhan, you've made this group a lot of fun over the past years, and I am really going to miss it.  You've been great colleagues and even better friends.

A big part of learning to do research and becoming a scientist is working with those who've come before you and learning everything they can teach you.  Andrzej, Roxanne, Jonathan, and Eric, I'm really grateful for everything you've taught me.  Andrzej, you've given me the best career advice I think I've ever gotten with ``Go learn physics.'' Challenging to do, but you've steered me well.

Flavio and Ornella, you've taught me more about neutrino physics and performing research than I can recount.  I look forward to many more years of you teaching me, and I hope you realize just how big an impact you've had on me and everyone around you.  Over the last year, Ornella, no one has guided me more in completing my work than you.

When I was looking for a research group during my first year of graduate school, a recently graduated student told me that working with Bonnie was ``the best decision I made during graduate school.''  Now, at the end, I can echo that statement.  Bonnie, you've treated me as a colleague from day one, always giving me respect and kindness while teaching me how to do physics. You've taught me how to be a better physicist, and how to succeed in science, and you set the example of how to treat everyone around you so well.

I know I thanked you first, Shannon, but it's worth doing again.  Thanks for helping me through every single day of graduate school.  I should amend my earlier statement: marrying you was the best decision I made in graduate school, or ever.  Thanks for everything.
